Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-05-15 second-draft/pnas-word-order-si.tex
SI

>>>>

\section{Efficiency and Prior Formalizations}

Here, we discuss how our formalization of efficiency relates to prior work testing functional explanations in linguistics.

%The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
%Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication.

%Recall our formalization of efficiency:
%\begin{equation}
%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
%\end{equation}
%\begin{equation}
%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
%\end{equation}
%\begin{equation}\label{eq:efficiency}
%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
%\end{equation}

Our objective function is identical to that of \cite{ferreri2003least}, and very similar to the functions in other computational work formalizing Zipf's ideas~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.

Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.

All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences communicated by the speaker to the listener.
By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
The syntactic structure (including the labels on the arcs) specifies how the meanings of words are combined.
Therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
In \cite{ferreri2003least, kemp2012kinship}, forms are words referring to objects.
In \cite{regier2015word} TODO
In \cite{zaslavsky2018efficient}, forms are color words, and meanings are TODO.

The accounts differ in the formalizations of complexity and diversity:
Like our model, \cite{ferreri2003least} model the complexity as the \emph{entropy} of the distribution of forms.
Closely related is the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., color words) and meanings (e.g., color chips).
This mutual information formalization is appropriate to encodes that are nondeterministic \cite{zaslavsky2018efficient}.
As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoded is deterministic, and which uses the entropy as the complexity measure.

A few models have used other measures that are not direct information-theoretic quantities.
In \cite{regier2015wordTODO}, the cost function is the number of different forms; in \cite{kemp2012kinship} it is the complexity of the concepts encoded.

%In terms of information theory \cite{shannon1948mathematical}, this corresponds to the notion of the rate-distortion tradeoff.
%In classical rate-distortion theory, the cost of a code is its rate, the mutual information between codes and underlying messages (as in the Information Bottleneck model).
% classical models of distortion are the Hamming distortion and 
%TODO also mention rate-distortion theory

- our function

similar: RSA, there it's utterance-level

\cite{frank2012predicting,goodman2013knowledge}

%- number of different forms 
%
%- complexity of the mapping 
%
%- MI between
%
%
%
%We model the first term as the 
% degree to which listeners can reconstruct dependency parse structures from an utterance, i.e., the \key{parseability} if the language. This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
%\begin{equation}
%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
%\end{equation}
%where the sum runs over all possible pairs of sentences $l$ and syntactic structures $t$ in the language.
%%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
%
%
%
%We formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
%In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
%Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
%%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
%
%In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
%\begin{equation}
%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
%\end{equation}
%where the sum runs over all possible sentences $u$ that belong to the language.
%%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. 
%
%
%
%The efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
%\begin{equation}\label{eq:efficiency}
%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
%\end{equation}
%with an interpolation weight $\lambda \in [0,1)$.
%In all results from Study 2 reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
%See SI appendix section 5 for mathematical discussion of $\lambda$, and robustness to other choices.
%
%
